# App-packages {#sec-app-pkgs .unnumbered}

```{r}
#| eval: true 
#| echo: false 
#| include: false
source("_common.R")
library(shinypak)
```

Expand the callout boxes below to review what we've covered in the last four chapters:

::: {.callout-note collapse='true'}

## [Whole app game (Chapter 1)]{style='font-size: 1.15em;'}

@sec-whole-game walks through the 'whole game' of the app-package development workflow. The application in this example comes from the [Packages](https://mastering-shiny.org/scaling-packaging.html) chapter of Mastering Shiny. 

- @sec-use-description
- @sec-use-data
- @sec-use-dependencies 
- @sec-use-r-code 
- @sec-use-load-all  
- @sec-use-package-tests  
- @sec-use-app-dot-r
    - @sec-use-pkg-doc-file
    - @sec-use-build-ignore
- @sec-use-license-file
- @sec-use-roxygen2 
    - @sec-use-namespace
- @sec-use-rstudio-project-options
- @sec-use-git 
- @sec-use-keyboard-shortcuts
- @sec-use-pkg-install 
- @sec-use-additional-files
    - @sec-use-readme-rmd
    - @sec-use-news-md
    - @sec-use-vignette
    

:::

::: {.callout-note collapse='true'}

## [Shiny (Chapter 2)]{style='font-size: 1.15em;'}

The [Shiny](shiny.qmd) chapter covers:

- @sec-shiny-projects
- @sec-shiny-folders
- @sec-shiny-files
    - @sec-shiny-modules 
    - @sec-shiny-standalone-app-fun
- @sec-adv-shiny-projects

:::

::: {.callout-note collapse='true'}

## [Packages (Chapter 3)]{style='font-size: 1.15em;'}

[Packages](packages.qmd) illustrates what separates R projects from R packages:

- @sec-packages-r-pkgs
    - @sec-packages-essential-ingredients 
    - @sec-packages-rstudio-build-pane
    - @sec-packages-positron-workspaces
- @sec-packages-app-pkgs
    - @sec-create-new-app-packages 
    - @sec-convert-shiny-app-packages 

:::

::: {.callout-note collapse='true'}

## [Development (Chapter 4)]{style='font-size: 1.15em;'}

@sec-dev introduced the `devtools` package and its core functions:

1. @sec-dev-load
2. @sec-dev-document and 
3. @sec-dev-install

:::

## App-packages

In the following chapters, we will continue developing our Shiny application within the R package structure. Below is a preview of each topic covered in the upcoming chapters.

### Documentation (Chapter 5)

@sec-doc covers documenting the code in your app-package with `roxygen2`:

- Required tags: 
    - @sec-roxygen2-md-support
    - @sec-roxygen2-title 
    - @sec-roxygen2-param-return and 
    - @sec-roxygen2-examples  
  
- App-package tags for modules and standalone app functions:
    - @sec-roxygen2-seealso  
    - @sec-roxygen2-family  
    - @sec-roxygen2-usage, and   
    - @sec-roxygen2-section

### Dependencies (Chapter 6)

@sec-depends dives into the external (i.e. add-on) functions and packages you've used to build your app-package. We'll also cover how to export functions and objects from your app-packages for others to use.

- @sec-depends-exports   
- @sec-depends-imports   

### Data (Chapter 7)

@sec-data covers how to include datasets in your app-package namespace, and the various kinds of data in R packages. 

- Storing and Documenting data:  
    - @sec-data-data 
    - @sec-document-data  
- Raw data files:   
    - @sec-data-data-raw   
- Other locations for data:   
    - @sec-data-inst-extdata    

### Launch (Chapter 8)

The @sec-launch chapter covers how to 1) run your application during development, 2) how to run your app using a standalone app function, and 3) how to deploy your application to shinyapps.io.

- App run functions: 
  -   @sec-launch-shiny-app
  -   @sec-launch-shiny-app-dir
  -   @sec-launch-run-app

- Launching app files: 
  -   @sec-launch-app-dot-r

### Resources (Chapter 9)

@sec-resources covers how to include external resources (i.e. the images, CSS, html, or other files that are typically included in the `www/` folder and served when you run your Shiny app). This chapter also covers other uses for the `isnt/` folder in app-packages.

- Static files: 
  -   @sec-resources-image-files  
  -   @sec-resources-add-resource-path  
- Data files:
  -   @sec-resources-inst-dev   
- Quarto apps:
  -   @sec-resources-inst-quarto   
- Production:
  -   @sec-resources-inst-prod


```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```
