# Code {#sec-code .unnumbered}

```{r}
#| eval: true 
#| echo: false 
#| include: false
source("_common.R")
library(shinypak)
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "y", size = "0.90", hsize = "0.95",
  look = "minimal",
  header = "Status",
  contents = "This section is under review.",
  fold = FALSE
)
```

## Review of Deploy

Let's briefly review what we've covered in the last three chapters:

### Docker (Chapter 15)

:::: {.callout-tip collapse='true' appearance='simple'}

#### [Chapter 15: GitHub Branches]{style='font-weight: bold; font-size: 1.15em;'}

::: {style='font-size: 0.95em; color: #282b2d;'}

The applications in Chapter 15 (Docker) can be accessed with the `launch()` or `get()` functions from the [`shinypak` R package](https://mjfrigaard.github.io/shinypak/):

```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| eval: false
# install.packages('pak')
pak::pak('mjfrigaard/shinypak')
library(shinypak)
```

Chapter 11 applications:
  
```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| collapse: true
list_apps(regex = '^15')
```

::: 

::::


### CI/CD (Chapter 16)

:::: {.callout-tip collapse='true' appearance='simple'}

#### [Chapter 16: GitHub Branches]{style='font-weight: bold; font-size: 1.15em;'}

::: {style='font-size: 0.95em; color: #282b2d;'}

The applications in Chapter 16 (CI/CD) can be accessed with the `launch()` or `get()` functions from the [`shinypak` R package](https://mjfrigaard.github.io/shinypak/):

```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| eval: false
# install.packages('pak')
pak::pak('mjfrigaard/shinypak')
library(shinypak)
```

Chapter 11 applications:
  
```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| collapse: true
list_apps(regex = '^16')
```

::: 

::::

### `pkgdown` (Chapter 17)

:::: {.callout-tip collapse='true' appearance='simple'}

#### [Chapter 17: GitHub Branches]{style='font-weight: bold; font-size: 1.15em;'}

::: {style='font-size: 0.95em; color: #282b2d;'}

The applications in Chapter 17 (`pkgdown`) can be accessed with the `launch()` or `get()` functions from the [`shinypak` R package](https://mjfrigaard.github.io/shinypak/):

```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| eval: false
# install.packages('pak')
pak::pak('mjfrigaard/shinypak')
library(shinypak)
```

Chapter 11 applications:
  
```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| collapse: true
list_apps(regex = '^17')
```

::: 

::::

## Code

### CSS

In [CSS](css.qmd), we'll discuss

### JavaScript

[JavaScript](js.qmd) covers 

### Python

The [Python](python.qmd) chapter covers 

## Style 
