# CSS {#sec-css}

---

[Dress Up Your App]{style="font-size: 1.40em; font-style: italic; color: #02577A;"}

```{r}
#| eval: true 
#| echo: false 
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "y", 
  header = "Caution",
  contents = "The contents for section are being revised. Thank you for your patience."
)
```

---

Cascading Style Sheets (CSS) is the stylesheet language for describing the presentation of HTML or XML documents. CSS defines how 'elements' should be displayed (layouts, colors, fonts, etc.). In Shiny app-packages, CSS can play a crucial role in customizing the appearance and enhancing the UI of an applications.



:::: {.callout-tip collapse='true' appearance='simple'}

## [Accessing applications]{style='font-weight: bold; font-size: 1.15em;'}

::: {style='font-size: 0.95em; color: #282b2d;'}

I've created the [`shinypak` R package](https://mjfrigaard.github.io/shinypak/) In an effort to make each section accessible and easy to follow:
  
Install `shinypak` using `pak` (or `remotes`):

```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| eval: false
# install.packages('pak')
pak::pak('mjfrigaard/shinypak')
```

Review the chapters in each section:
  
```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| collapse: true
library(shinypak)
list_apps(regex = 'css')
```

Launch an app with `launch()`

```{r}
#| code-fold: false 
#| eval: false
launch(app = "18_css")
```

Download an app with `get_app()`

```{r}
#| code-fold: false 
#| eval: false
get_app(app = "18_css")
```

::: 

::::

<!--
https://www.w3schools.com/css/css_intro.asp
-->
