# Deploy {#sec-deploy .unnumbered}

```{r}
#| eval: true 
#| echo: false 
#| include: false
source("_common.R")
library(shinypak)
```

Expand the callout boxes below to review the previous testing chapters:

:::: {.callout-tip collapse="true" title="Review Chapters 14 - 18"}

::: {.callout-note collapse='true'}

#### Test suite (Chapter 14)

[Test suite](test_suite.qmd) covers:

- @sec-tests-suite-use-testthat   
- @sec-tests-suite-bdd-intro  
- @sec-tests-suite-keyboard-shortcuts 

:::

::: {.callout-note collapse='true'}

#### Specifications (Chapter 15)

[Specifications](test_specs.qmd) dives into:

- @sec-tests-specs-app-reqs   
    - @sec-tests-specs-user-specs   
    - @sec-tests-specs-features   
    - @sec-tests-specs-func-reqs      
- @sec-tests-specs-bdd-testthat   
- @sec-tests-specs-traceability-matrix    

:::: {.callout-tip collapse='true' appearance='default'}

#### Chapter 15 Applications

::: {style='font-size: 0.95em; color: #282b2d;'}

The applications in Chapter 15 (Specifications) can be accessed with the `launch()` or `get()` functions from the [`shinypak` R package](https://mjfrigaard.github.io/shinypak/):

```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| eval: false
# install.packages('pak')
pak::pak('mjfrigaard/shinypak')
library(shinypak)
```

Chapter 15 applications:
  
```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| collapse: true
list_apps(regex = '^15')
```

::: 

::::

:::

::: {.callout-note collapse='true'}

#### Test tools (Chapter 16)

[Test tools](test_tools.qmd) covers: 

- @sec-test-fixtures  
- @sec-test-helpers   
- @sec-test-coverage-active-file    
- @sec-test-snaps-review    

:::: {.callout-tip collapse='true' appearance='default'}

#### Chapter 16 Applications

::: {style='font-size: 0.95em; color: #282b2d;'}

The applications in Chapter 16 (Test tools) can be accessed with the `launch()` or `get()` functions from the [`shinypak` R package](https://mjfrigaard.github.io/shinypak/):

```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| eval: false
# install.packages('pak')
pak::pak('mjfrigaard/shinypak')
library(shinypak)
```

Chapter 16 applications:
  
```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| collapse: true
list_apps(regex = '^16')
```

::: 

::::

:::

::: {.callout-note collapse='true'}

#### Testing modules (Chapter 17)

[Testing modules](test_modules.qmd) includes strategies for: 

- @sec-tests-mods-returned    
- @sec-tests-mods-args-list     
- @sec-tests-mods-coverage    
  - @sec-tests-mods-outputs   

:::: {.callout-tip collapse='true' appearance='default'}

#### Chapter 17 Applications

::: {style='font-size: 0.95em; color: #282b2d;'}

The applications in Chapter 17 (Testing modules) can be accessed with the `launch()` or `get()` functions from the [`shinypak` R package](https://mjfrigaard.github.io/shinypak/):

```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| eval: false
# install.packages('pak')
pak::pak('mjfrigaard/shinypak')
library(shinypak)
```

Chapter 17 applications:
  
```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| collapse: true
list_apps(regex = '^17')
```

::: 

::::

:::

::: {.callout-note collapse='true'}

#### System tests (Chapter 18)

- @sec-tests-system-shinytest2    
- @sec-tests-system-recording-tests   
- @sec-tests-system-test-mode   

:::: {.callout-tip collapse='true' appearance='default'}

#### Chapter 18 Applications

::: {style='font-size: 0.95em; color: #282b2d;'}

The applications in Chapter 18 (System tests) can be accessed with the `launch()` or `get()` functions from the [`shinypak` R package](https://mjfrigaard.github.io/shinypak/):

```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| eval: false
# install.packages('pak')
pak::pak('mjfrigaard/shinypak')
library(shinypak)
```

Chapter 18 applications:
  
```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| collapse: true
list_apps(regex = '^18')
```

::: 

::::

:::

::::

# Deploy

This section covers two common methods for deploying a Shiny app-package: 1) Docker images and containers, 2) using GitHub Actions for Continuous Integration (CI) and Continuous Deployment (CD), and 3) building and deploying a package website. 

## Docker

[Docker](docker.qmd) covers how using Docker can simplify setting up your R environment and reduce the "it works on my machine" problem:
  
-   @sec-docker-what-is-docker  
-   @sec-docker-install   
-   @sec-docker-dockerfile    
-   @sec-docker-build-image   
-   @sec-docker-run-container   
-   @sec-docker-golem   
-   @sec-docker-app-packages    
  
## GitHub Actions

Continuous Integration/Continuous Deployment (CI/CD) can be implemented for your Shiny app-package using [GitHub actions](gha.qmd). This chapter covers the following workflows:

-   @sec-gha-style    
-   @sec-gha-shiny    
-   @sec-gha-shiny-docker   

## Package websites 

[`pkgdown`](pkgdown.qmd) helps us easily create websites for our app-packages. This chapter covers:
  
-   @sec-pkgdown-setup  
-   @sec-pkgdown-site-contents    
-   @sec-pkgdown-layouts    
-   @sec-pkgdown-reference    
-   @sec-pkgdown-workflow   