# Glossary of terms {.unnumbered}

```{r}
#| eval: true 
#| echo: false 
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "r",
  header = "ALERT!",
  contents = "This section is currently under development. Thank you for your patience."
)
```

## Shiny apps

### Utility/helper functions 

### External resources

### Module 

### Standalone app function

## Packages 

### Project 

Any directory of [R executable files](https://cran.r-project.org/) with an RStudio/Posit workbench project file (`.Rproj`).

### Package

A directory of functions, documentation, and/or data can be installed and loaded into an R session. Packages include the necessary dependency management (`NAMESPACE`) and metadata files (`DESCRIPTION`) *and* has access to the package development tools in RStudio/Posit Workbench.

### App-package

An app-package is a package containing a shiny application. App-packages have all of the functionality of a standard R package, but also contain the files and folders required to successfully develop, run, and deploy a shiny application.

## [`devtools`]{style="font-size: 1.05em;"}

end `glossary.qmd`