# Introduction {#sec-intro .unnumbered}

```{r}
#| eval: true 
#| echo: false 
#| include: false
source("_common.R")
"#3a506b"
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "o", look = "minimal",
  header = "Caution",
  contents = "This section is still being developed. The contents are subject to change.",
  fold = FALSE
)
```

A guiding principle throughout this book is,

[Organize your Shiny app code before you share it with others.]{style='font-size: 1.15em; font-style: italic; font-weight: bold; color: #3a506b;'}

I could reduce this statement to ‘organize your code,’ but it's especially true for Shiny applications destined for deployment in a production environment. In R, the standardized and widely accepted way to organize code as an R package.[^intro-r-pkgs] 

```{r}
#| label: co_box_production
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "b",
  header = "What does it mean to 'put something into production?'",
  contents = "
  
‘Production’ usually means passing the code from your personal development environment into your company's cloud-based server environment, which typically involves bundling your app in a structure that can be shared, installed, tested, and launched.
  
> '*I think the easiest way to think about it for me is that we develop a model in one computational environment--think of this as maybe your laptop or maybe you work in a server environment--still, it's in one place, and it turns out the software that you need to have installed there is about tuning, training, etc. **Putting something into production is getting it out of that computational environment and successfully carrying it over to a new computational environment.** For many people, this might be like a  cloud computing environment. It might be some kind of server your organization has, and we need to take it, lift it, and then successfully have it working.*' - Julia Silge, [What is 'production' anyway? MLOps for the curious (SatRdays London 2023)]((https://youtu.be/53-x5OFMGjA?t=1411))
  
I've added emphasis and edited this for clarity.
  
  ",
  fold = TRUE, 
  look = "minimal"
)
```

[^intro-r-pkgs]: David Neuzerling has a great [post](https://mdneuzerling.com/post/data-science-workflows/) on the benefits of structuring your project as a package 

## Packages help you

The first and obvious benefit to structuring your Shiny app as an R package is that it simplifies file and folder management. If every Shiny app project you develop is structured as an R package, it removes the time you spend manually creating directories (or re-orienting yourself to each project's structure). 

If you're using Posit Workbench, the R package structure will give you access to a well-designed IDE for Shiny applications. Posit Workbench has tools to help develop and debug functions, create and run unit tests, store internal and external data, manage dependencies, and write help files and long-form documentation. 

## Packages help them

> "*a package [is] the fundamental unit of shareable, reusable, and reproducible R code.*" - [R Packages, 2ed](https://r-pkgs.org/)

It's safe to assume the code used to build Shiny apps being deployed to a production environment will be seen (and hopefully used) by others. R packages make sharing your hard work with your colleagues easier because it removes their need to figure out where everything is, how it all fits together, and how it all (hopefully) works.

## Scalable 

[Great R packages define and solve common problems.]{style='font-size: 1.15em; font-style: italic; font-weight: bold; color: #3a506b;'} 

Suppose you use R to perform analyses, design data visualizations, or build and run reports. If you currently use `source()` to load any utility functions to perform this work, I suggest putting those functions in a package. Doing this will help extend your mental model from the specific use cases (i.e., "*X code performs task Y*") to a model for their more general uses (i.e., "*X package performs tasks like Y*").

The beauty of an R package mental model is that you’ll inevitably notice the similarities across common problems. Creating packages that define and solve common problems in your workflow can sometimes be some of the most popular/valuable contributions (see [`datapasta`](https://milesmcbain.github.io/datapasta/) and [`reprex`](https://reprex.tidyverse.org/)).


## How to read this book 

The chapters in this book (roughly) represent the steps of R package development, but from the lens of an existing shiny application.[^intro-bwas] Each topic can (and should) be applied when creating new app-packages. Still, in my experience, many Shiny developers have existing applications they'd like to convert into an R package.

The first four chapters focus on a shiny app project:

-   @sec-whole-game is a 'TLDR' for the development of a toy app-package. This chapter gives a high-level overview of the package development workflow.[^r-pkgs-whole-game]

-   In @sec-shiny we'll dive into shiny development, focusing on the files and folders found in most shiny applications.

-   R packages will be introduced in @sec-pkgs, and we'll cover the differences between shiny app projects, R packages, and shiny app-packages. 

-   @sec-dev introduces `devtools` and the app-package development workflow.

[^r-pkgs-whole-game]: R Packages, 2ed, has a [similar chapter](https://r-pkgs.org/whole-game.html) that covers developing standard R packages. 

[^intro-bwas]: The original code and data for the application in this book comes from the [Building Web Applications with Shiny (BWAS)](https://rstudio-education.github.io/shiny-course/) course.

Transitioning from programming in R to building Shiny applications and writing R packages is complicated. It involves a shift in focus, skill set, and development practices. Reactivity and modules are a steep learning curve, especially if you're also trying to understand how to develop an R package (with an application in it!).[^intro-golem] [^intro-mastering-shiny]

[^intro-golem]: [Engineering Production-Grade Shiny Apps](https://engineering-shiny.org/) and the [`golem` package](https://thinkr-open.github.io/golem/) assumes readers "*are comfortable with building an R package.*"

[^intro-mastering-shiny]: Mastering Shiny includes an example for  *[converting](https://mastering-shiny.org/scaling-packaging.html?#converting-an-existing-app)* an existing application with `usethis::use_description()`, but not creating a new app-package (i.e., with `usethis::create_package()`).

The chapters of this book are intended to help with the package development end of that learning curve. 

### Code examples

```{r}
#| label: git_box_main_demo
#| echo: false
#| results: asis
#| eval: true
git_margin_box(
  fig_pw = '75%', 
  branch = "main", 
  repo = 'moviesApp')
```


The code files for each chapter are stored in GitHub repositories. Major code changes are stored in branches. Whenever a new branch or repo is used, you'll see the **Git Branch** icon with a link to the repo in the margin.



<!--
:::: {.callout-important collapse="true" appearance='simple'}

### [`monthApp`]{style='font-size: 1.10em;'} from Mastering Shiny

::: {style='font-size: 1.05em; color: #696969;'}

If you happened to download or clone the [`monthApp`](https://github.com/hadley/monthApp) example from the [Packages Chapter](https://mastering-shiny.org/scaling-packaging.html), you may have notice a few things:[^pkgs-0]

1. `DESCRIPTION` contains additional fields that are not addressed.[^pkgs-1] 

2. `monthApp.Rproj` has been configured to work with package.[^pkgs-2a]

3. `devtools::build()` ([<kbd>Ctrl/Cmd</kbd> + <kbd>Shift</kbd> + <kbd>B</kbd>]{style="font-weight: bold; font-size: 0.90em"}) has been configured to have additional behaviors.[^pkgs-2b]

4. Dependency management is discussed briefly but not documented in the application.[^pkgs-3]
  
5. `monthApp` doesn't have any help files[^pkgs-4] or tests.[^pkgs-5] 


[^pkgs-0]: The [Packages](https://mastering-shiny.org/scaling-packaging.html) chapter from [Mastering Shiny](https://mastering-shiny.org/) gets your '*toes into the water of package development*', but in my opinion, is probably not enough for you to get your shiny app project into an R package that's ready to be shipped into most production environments.

[^pkgs-1]: The [`DESCRIPTION`](https://github.com/hadley/monthApp/blob/master/DESCRIPTION) file in `monthApp` contains fields not included in the arguments passed to `usethis::use_description()` (these were added [separately](https://github.com/hadley/monthApp/commit/9899da4c89f4114a0611b415f4a329daeb753069)), and a few arguments can be adapted or removed to save time.

[^pkgs-2a]: The `.Rproj` file is the connection between the `DESCRIPTION` fields, the code in the `R/` folder, and the IDE's **Build** pane or “package development mode.” If you already have an `.Rproj` file in your shiny app project, you can activate these settings under **Tools** > **Project Options** > **Build Tools**. 

[^pkgs-2b]: The [`monthApp.Rproj`](https://github.com/hadley/monthApp/blob/9899da4c89f4114a0611b415f4a329daeb753069/monthApp.Rproj#L22) file contains additional settings you can add with **Tools** > **Project Options** > **Build Tools**, then under **Generate documentation with Roxygen**, click **Configure** and select **Install and Restart**.

[^pkgs-3]: The [`NAMESPACE` file](https://github.com/hadley/monthApp/blob/master/NAMESPACE) is empty and the `Imports` field is missing from the [`DESCRIPTION`](https://github.com/hadley/monthApp/blob/master/DESCRIPTION) (although the chapter discusses [importing `shiny`](https://mastering-shiny.org/scaling-packaging.html#r-cmd-check)).

[^pkgs-4]: `roxygen2` is mentioned, but it's beyond the scope of [Mastering Shiny](https://mastering-shiny.org/), so for now just know `roxygen2` syntax is placed in the code below `R/` to create the help files and the `NAMESPACE`.

[^pkgs-5]: `monthApp` doesn't have the testing infrastructure provided by [`testthat`](https://testthat.r-lib.org/), but this can be quickly created using `usethis::use_testthat()` (and an entire chapter is dedicated to [testing](https://mastering-shiny.org/scaling-testing.html).



::: 

::::

-->