# Shiny & LLMs {#sec-shiny-llm .unnumbered}

```{r}
#| eval: true 
#| echo: false 
#| include: false
source("_common.R")
library(shinypak)
```

Since I began writing this book (in late 2022), the number of large-language models (LLMs) has grown significantly. The chapters in this section introduce a few of the popular LLM tools available for building Shiny apps and R packages. Given these tools' rapidly evolving landscape and nature, I expect these chapters to change frequently. Please open a GitHub issue if there is anything outdated, incorrect, or missing.

:::: {.callout-tip collapse="true" title="Review Chapters 19 - 22"}

::: {.callout-note collapse='true'}

#### shinyapps.io (Chapter 19)

- @sec-shinyappsio-publish    
- @sec-shinyappsio-deploy-log   
- @sec-shinyappsio-rsconnect-dir    
- @sec-shinyappsio-app-logs   

:::: {.callout-tip collapse='true' appearance='default'}

#### Chapter 19 Applications

::: {style='font-size: 0.95em; color: #282b2d;'}

The applications in Chapter 19 (shinyapps.io) can be accessed with the `launch()` or `get()` functions from the [`shinypak` R package](https://mjfrigaard.github.io/shinypak/):

```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| eval: false
# install.packages('pak')
pak::pak('mjfrigaard/shinypak')
library(shinypak)
```

Chapter 19 applications:
  
```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| collapse: true
list_apps(regex = '^19')
```

::: 

::::

:::

::: {.callout-note collapse='true'}

#### Docker (Chapter 20)

[Docker](docker.qmd) dives into:

- @sec-docker-what-is-docker  
- @sec-docker-install     
- @sec-docker-dockerfile    
- @sec-docker-build-image   
- @sec-docker-run-container   
- @sec-docker-golem   
- @sec-docker-app-packages    

:::: {.callout-tip collapse='true' appearance='default'}

#### Chapter 20 Applications

::: {style='font-size: 0.95em; color: #282b2d;'}

The applications in Chapter 20 (Docker) can be accessed with the `launch()` or `get()` functions from the [`shinypak` R package](https://mjfrigaard.github.io/shinypak/):


```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| eval: false
# install.packages('pak')
pak::pak('mjfrigaard/shinypak')
library(shinypak)
```

Chapter 20 applications:
  
```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| collapse: true
list_apps(regex = '^20')
```

::: 

::::

:::

::: {.callout-note collapse='true'}

#### GitHub Actions (Chapter 21)

The [GitHub Actions](gha.qmd) chapter includes three examples of CI/CD workflows:

- @sec-gha-style    
- @sec-gha-shiny    
- @sec-gha-shiny-docker   

:::: {.callout-tip collapse='true' appearance='default'}

#### Chapter 21 Applications

::: {style='font-size: 0.95em; color: #282b2d;'}

The applications in Chapter 21 (GitHub Actions) can be accessed with the `launch()` or `get()` functions from the [`shinypak` R package](https://mjfrigaard.github.io/shinypak/):


```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| eval: false
# install.packages('pak')
pak::pak('mjfrigaard/shinypak')
library(shinypak)
```

Chapter 21 applications:
  
```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| collapse: true
list_apps(regex = '^21')
```

::: 

::::

:::

::: {.callout-note collapse='true'}

#### pkgdown (Chapter 22)

[pkgdown](pkgdown.qmd) describes:

- @sec-pkgdown-setup    
- @sec-pkgdown-site-contents            
- @sec-pkgdown-layouts      
    - @sec-pkgdown-reference      
- @sec-pkgdown-workflow       
- @sec-pkgdown-github     

:::: {.callout-tip collapse='true' appearance='default'}

#### Chapter 22 Applications

::: {style='font-size: 0.95em; color: #282b2d;'}

The applications in Chapter 22 (`pkgdown`) can be accessed with the `launch()` or `get()` functions from the [`shinypak` R package](https://mjfrigaard.github.io/shinypak/):


```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| eval: false
# install.packages('pak')
pak::pak('mjfrigaard/shinypak')
library(shinypak)
```

Chapter 22 applications:
  
```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| collapse: true
list_apps(regex = '^22')
```

::: 

::::

:::


::::

## Shiny Assistant

The [Shiny Assistant](https://gallery.shinyapps.io/assistant/) covers: 

> "*Think of AI tools as a friend who is a know-it-all, over-confident, presumptuous, and lacks self-awareness. Despite these imperfections, they remain your friend.*"

## shinychat 

## ellmer

## chores 