# rhino (CI/CD) {#sec-rhino-appendix}

```{r}
#| eval: true 
#| echo: false 
#| include: false
source("_common.R")
library(testthat)
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "o", 
  look = "minimal",
  header = "Alert",
  contents = "The contents for section are being developed. Thank you for your patience."
)
```

## Rhino CI/CD

The `rhino` framework includes the following GitHub Actions workflow file with new apps: 

<details>

<summary>.github/workflows/rhino-test.yml</summary>

```yaml
name: Rhino branch test, 20_rhino
name: Rhino Test
on: push
permissions:
  contents: read
  
jobs:
  main:
    name: Run linters and tests
    runs-on: ubuntu-20.04
    env:
      GITHUB_PAT: ${{ secrets.GITHUB_TOKEN }}
    steps:
      - name: Checkout repo
        uses: actions/checkout@v2

      - name: Extract R version from lockfile
        run: printf 'R_VERSION=%s\n' "$(jq --raw-output .R.Version renv.lock)" >> $GITHUB_ENV

      - name: Setup R
        uses: r-lib/actions/setup-r@v2
        with:
          r-version: ${{ env.R_VERSION }}

      - name: Setup system dependencies
        run: >
          sudo apt-get update && sudo apt-get install --yes
          libcurl4-openssl-dev

      - name: Restore renv from cache
        uses: actions/cache@v2
        env:
          CACHE_KEY: renv-${{ runner.arch }}-${{ runner.os }}-${{ env.R_VERSION }}
        with:
          path: renv/library
          key: ${{ env.CACHE_KEY }}-${{ hashFiles('renv.lock') }}
          restore-keys: ${{ env.CACHE_KEY }}-

      - name: Sync renv with lockfile
        shell: Rscript {0}
        run: |
          options(renv.config.cache.symlinks = FALSE)
          renv::restore(clean = TRUE)

      - name: Setup Node
        uses: actions/setup-node@v2
        with:
          node-version: 16

      - name: Lint R
        if: always()
        shell: Rscript {0}
        run: rhino::lint_r()

      - name: Lint JavaScript
        if: always()
        shell: Rscript {0}
        run: rhino::lint_js()

      - name: Lint Sass
        if: always()
        shell: Rscript {0}
        run: rhino::lint_sass()

      - name: Build JavaScript
        if: always()
        shell: Rscript {0}
        run: rhino::build_js()

      - name: Build Sass
        if: always()
        shell: Rscript {0}
        run: rhino::build_sass()

      - name: Run R unit tests
        if: always()
        shell: Rscript {0}
        run: rhino::test_r()

      - name: Run Cypress end-to-end tests
        if: always()
        uses: cypress-io/github-action@v5
        with:
          working-directory: .rhino # Created by earlier commands which use Node.js
          start: npm run run-app
          project: ../tests
          wait-on: 'http://localhost:3333/'
          wait-on-timeout: 60
```

</details>

The sections are described below:

### Trigger 

The `.github/workflows/rhino-test.yml` file runs on pushes to the repo containing a `rhino` app.
   
```yml
name: Rhino Test
on: push
```

### Permissions

The workflow only has `read` permission for the repository. Read more about setting permissions [here](https://docs.github.com/en/actions/using-jobs/assigning-permissions-to-jobs#defining-access-for-the-github_token-scopes).

```yml
permissions:
  contents: read
```

### Jobs

`rhino-test.yml` contains a single job named `main` with the following `steps`:

```yml
jobs:
  main:
    name: Run linters and tests
    runs-on: ubuntu-20.04
    env:
      R_VERSION: '4.1.0'
      GITHUB_PAT: ${{ secrets.GITHUB_TOKEN }}
```


#### Step: Checkout Code, R version, Dependencies

Checks out the code using `actions/checkout@v2`, extracts the R version from a lockfile (`renv.lock`) and sets it as an environment variable (`env.R_VERSION`), and uses `apt-get` to install system dependencie (`libcurl4-openssl-dev`)

```yml
    steps:
      - name: Checkout repo
        uses: actions/checkout@v2

      - name: Setup R
        uses: r-lib/actions/setup-r@v1
        with:
          r-version: ${{ env.R_VERSION }}

      - name: Setup system dependencies
        run: >
          sudo apt-get update && sudo apt-get install --yes
          libcurl4-openssl-dev
```

#### Step: Restore and sync `renv`

The R dependencies in `renv.lock` are restored from cache with `actions/cache@v2`, then the R environment is synchronized with the lockfile (ensuring all necessary R packages are installed).

```yml

      - name: Restore renv from cache
        uses: actions/cache@v2
        env:
          CACHE_KEY: renv-${{ runner.arch }}-${{ runner.os }}-${{ env.R_VERSION }}
        with:
          path: renv/library
          key: ${{ env.CACHE_KEY }}-${{ hashFiles('renv.lock') }}
          restore-keys: ${{ env.CACHE_KEY }}-
          
      - name: Sync renv with lockfile
        shell: Rscript {0}
        run: |
          options(renv.config.cache.symlinks = FALSE)
          renv::restore(clean = TRUE)
```

#### Step: Node.js

The Node.js environment is set up with version 16.

```yml
      - name: Setup Node
        uses: actions/setup-node@v2
        with:
          node-version: 16
```

#### Step: Linters

Lints R (`rhino::lint_r()`), JavaScript (`rhino::lint_js()`), and Sass (`rhino::lint_sass()`) code for quality and style consistency.

```yml
      - name: Lint R
        if: always()
        shell: Rscript {0}
        run: rhino::lint_r()

      - name: Lint JavaScript
        if: always()
        shell: Rscript {0}
        run: rhino::lint_js()

      - name: Lint Sass
        if: always()
        shell: Rscript {0}
        run: rhino::lint_sass()
```

#### Step: Unit Tests

Executes `testthat` unit tests written in R (`rhino::test_r()`)


```yml
      - name: Run R unit tests
        if: always()
        shell: Rscript {0}
        run: rhino::test_r()
```

#### Step: Cypress End-to-End Tests

Specifies the working directory `.rhino/`, starting command for the application (`npm run run-app`), project directory for tests (`../tests`), URL to wait for (`'http://localhost:3333/'`), and a timeout for waiting (`60`) for Cypress to run end-to-end tests.

```yml
      - name: Run Cypress end-to-end tests
        if: always()
        uses: cypress-io/github-action@v5
        with:
          working-directory: .rhino # Created by earlier commands which use Node.js
          start: npm run run-app
          project: ../tests
          wait-on: 'http://localhost:3333/'
          wait-on-timeout: 60

```
