# Special topics {#sec-spec-topics .unnumbered}

```{r}
#| eval: true 
#| echo: false 
#| include: false
source("_common.R")
# pak::pak('mjfrigaard/shinypak')
library(shinypak)
```

This section holds a collection of topics you might find useful when developing your app-package.

:::: {.callout-tip collapse="true" title="Review Chapters 26 - 29"}

::: {.callout-note collapse='true'}

#### Shiny Assistant (Chapter 26)

The [shiny_assist](shiny_assist.qmd) chapter includes a series of prompts for developing an app-package:

- @sec-shiny-assist-ui    
- @sec-shiny-assist-p1  
- @sec-shiny-assist-p2  
- @sec-shiny-assist-p3  
- @sec-shiny-assist-p4  
- @sec-shiny-assist-p5  
- @sec-shiny-assist-p6
- @sec-shiny-assist-finishing-touches    

:::: {.callout-tip collapse='true' appearance='default'}

#### Chapter 26 Applications

::: {style='font-size: 0.95em; color: #282b2d;'}

The applications in Chapter 26 can be accessed with the `launch()` or `get()` functions from the [`shinypak` R package](https://mjfrigaard.github.io/shinypak/):

```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| eval: false
# install.packages('pak')
pak::pak('mjfrigaard/shinypak')
library(shinypak)
```

Chapter 26 applications:
  
```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| collapse: true
list_apps(regex = '^26')
```

::: 

::::

:::

::: {.callout-note collapse='true'}

#### ellmer (Chapter 27)

[ellmer](ellmer.qmd) covers:

- @sec-ellmer-install
    - @sec-ellmer-chat    
    - @sec-ellmer-prompts      
- @sec-ellmer-shinydb   
    - @sec-ellmer-sys-prompt        
    - @sec-ellmer-update    
    - @sec-ellmer-refactor      
    - @sec-ellmer-overhaul    
- @sec-ellmer-finishing-touches       

:::: {.callout-tip collapse='true' appearance='default'}

#### Chapter 27 Applications

::: {style='font-size: 0.95em; color: #282b2d;'}

The applications in Chapter 27 (`ellmer`) can be accessed with the `launch()` or `get()` functions from the [`shinypak` R package](https://mjfrigaard.github.io/shinypak/):


```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| eval: false
# install.packages('pak')
pak::pak('mjfrigaard/shinypak')
library(shinypak)
```

Chapter 27 applications:
  
```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| collapse: true
list_apps(regex = '^27')
```

::: 

::::

:::

::: {.callout-note collapse='true'}

#### chores (Chapter 28)

The [chores](chores.qmd) chapter includes:

- @sec-chores-navbar-app    
    - @sec-chores-navbar-modules      
        - @sec-chores-compare-module    
        - @sec-chores-counts-module   
        - @sec-chores-dist-module   
        - @sec-chores-awards-module   
    - @sec-chores-navbar-utils    
        - @sec-chores-utils-name-case    
        - @sec-chores-utils-dist-var-summary     
        - @sec-chores-utils-create-movie-awards     
- @sec-shiny-chores-setup     
    - @sec-shiny-chores-addin     
- @sec-chores-helper-addin      
- @sec-chores-custom-helpers  
    - @sec-chores-ext-pkg

:::: {.callout-tip collapse='true' appearance='default'}

#### Chapter 28 Applications

::: {style='font-size: 0.95em; color: #282b2d;'}

The applications in Chapter 28 (`chores`) can be accessed with the `launch()` or `get()` functions from the [`shinypak` R package](https://mjfrigaard.github.io/shinypak/):


```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| eval: false
# install.packages('pak')
pak::pak('mjfrigaard/shinypak')
library(shinypak)
```

Chapter 28 applications:
  
```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| collapse: true
list_apps(regex = '^28')
```

::: 

::::

:::

::: {.callout-note collapse='true'}

### gander (Chapter 29)

The [gander](gander.qmd) chapter covers:

- @sec-gander-config
    - @sec-gander-shortcut    
- @sec-gander-navbar-app    
- @sec-gander-dev-addin   
    - @sec-gander-suffix    
- @sec-gander-context     
    - @sec-gander-context-system      
    - @sec-gander-context-user      
    - @sec-gander-context-assistant     
- @sec-gander-visualizations      
- @sec-gander-rmd-report    

:::: {.callout-tip collapse='true' appearance='default'}

#### Chapter 29 Applications

::: {style='font-size: 0.95em; color: #282b2d;'}

The applications in Chapter 29 (`gander`) can be accessed with the `launch()` or `get()` functions from the [`shinypak` R package](https://mjfrigaard.github.io/shinypak/):


```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| eval: false
# install.packages('pak')
pak::pak('mjfrigaard/shinypak')
library(shinypak)
```

Chapter 29 applications:
  
```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| collapse: true
list_apps(regex = '^29')
```

::: 

::::

:::

::::

## Shiny UI Editor 

[Shiny UI Editor](shiny_ui_edit.qmd) covers...

## Code style

[Code style](code.qmd) discusses...

## App data

In [App data](app_data.qmd)...

- @sec-reactiveValues   
    - @sec-testing-reactive-values     
- @sec-session-user-data    


## Dependency hell

The @sec-entanglement chapter


```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```
