# Special topics {#sec-spec-topics .unnumbered}

```{r}
#| eval: true 
#| echo: false 
#| include: false
source("_common.R")
```

```{r}
#| label: git_co_box
#| echo: false
#| results: asis
#| eval: true
git_co_box()
```

This section holds a collection of topics you might find useful when developing your app-package.

## Debugging

The [Debugging](debugging.qmd) chapter covers how to use `browser()` (see @sec-browser) and the IDE's debugger console to 'step through' a function line-by-line. This chapter also covers how to send and view `print()` statements in the UI (see @sec-verbatim-reactives) 


## App data

In [App data](app_data.qmd), @sec-reactiveValues demonstrates how we can use [`reactiveValues()`](https://shiny.posit.co/r/reference/shiny/0.11/reactivevalues) to store and retrieve reactive values from a 'persistent' object in our app. @sec-testing-reactive-values also has examples of testing modules with values from a `reactiveValues()` object.

@sec-session-user-data gives an example of how to store persistent, non-reactive objects in your application (see also @sec-external-inst-dev). 




## Stack traces 


The [Stack traces ](stack_traces.qmd) chapter...

## Entanglement

The [Entanglement](Entanglement.qmd) chapter...




