# Stack traces {#sec-stack-traces}

```{r}
#| eval: true 
#| echo: false 
#| include: false
source("_common.R")
```

```{r}
#| label: git_box_install
#| echo: false
#| results: asis
#| eval: true
git_margin_box(
  contents = "install")
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "y", 
  look = "minimal",
  header = "Alert",
  contents = "The contents for section are being revised. Thank you for your patience."
)
```

In this chapter covers how to read stack traces (or call stacks), which is the printout that follows programming errors. 

```{r}
#| label: git_box_27_stack-traces
#| echo: false
#| results: asis
#| eval: true
git_margin_box(
  contents = "launch",
  fig_pw = '75%', 
  branch = "27_stack-traces", 
  repo = 'moviesApp')
```

:::: {.callout-tip collapse='true' appearance='simple'}

## [Accessing applications]{style='font-weight: bold; font-size: 1.15em;'}

::: {style='font-size: 0.95em; color: #282b2d;'}

I've created the [`shinypak` R package](https://mjfrigaard.github.io/shinypak/) In an effort to make each section accessible and easy to follow:
  
Install `shinypak` using `pak` (or `remotes`):

```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| eval: false
# install.packages('pak')
pak::pak('mjfrigaard/shinypak')
```

Review the chapters in each section:
  
```{r}
#| code-fold: false 
#| message: false
#| warning: false
#| collapse: true
library(shinypak)
list_apps(regex = '^27')
```

Launch the app: 

```{r}
#| code-fold: false 
#| eval: false
launch(app = "27_stack-traces")
```

Download the app: 

```{r}
#| code-fold: false 
#| eval: false
get_app(app = "27_stack-traces")
```

::: 

::::

```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```